---
title: "ST 495 HW9"
author: "Eric Warren"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      fig.height = 7,
                      fig.width = 10)
```

## Honor Code {.tabset}

I have neither given nor received unauthorized assistance on this test or assignment. -**Eric Warren**

### Problem 1 {.tabset}

#### Part A

Here we are going to load in the `HairEyeColor` data set. Convert the 3-way table into a data frame with 32 rows and 4 columns. Partition the data frame into two parts --- one for male and one for female. For each gender, construct a stacked barplot of the frequencies with Hair color in the X-axis and Eye color in the Y-axis, with different eye colors as different stacks of the bars. Write down your interpretation of the plots.
```{r}
# Changing data into dataframe
library(tidyverse)
hairEyeColor <- as.data.frame(ftable(HairEyeColor))

# Partition the data
male <- filter(hairEyeColor, Sex == "Male")
female <- filter(hairEyeColor, Sex != "Male")

# Making a stacked barplot of males and females side by side
library(gridExtra)
plot1 <- male %>%
  ggplot(aes(x = Hair, y = Freq, fill = Eye)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(title = "Male Eye and Hair Colors",
       caption = "Eric Warren",
       x = "Hair Color",
       y = "Frequency",
       color = "Eye Color") +
  theme_bw() +
  theme(plot.title = element_text(color = "blue",
                                  face = "bold",
                                  size = 16,
                                  hjust = 0.5))
plot2 <- female %>%
  ggplot(aes(x = Hair, y = Freq, fill = Eye)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  labs(title = "Female Eye and Hair Colors",
       caption = "Eric Warren",
       x = "Hair Color",
       y = "Frequency",
       color = "Eye Color") +
  theme_bw() +
  theme(plot.title = element_text(color = "blue",
                                  face = "bold",
                                  size = 16,
                                  hjust = 0.5))
grid.arrange(plot1, plot2, ncol = 2)
```

As we can see from this plot, they are roughly the same except there are more blond hair, blue eye females than males. Also most blonds have blue eyes and most people with black hair have brown eyes. The other hair colors are toss ups and the hair color most prevalent is brown.

#### Part B

Now consider the original 3-way table. There are three features: Hair color, Eye color, and Sex. For each pair of features, carry out a chi-square test of independence and report whether there is an association between features.
```{r}
# Make our new 2 way tables
hairVsEye <- HairEyeColor[ , , 1] + HairEyeColor[ , , 2]
hairVsSex <- HairEyeColor[ , 1, ] + HairEyeColor[ , 2, ] + HairEyeColor[ , 3, ] + HairEyeColor[ , 4, ]
eyeVsSex <- HairEyeColor[ 1, , ] + HairEyeColor[2, , ] + HairEyeColor[3, , ] + HairEyeColor[4, , ]

# Do Chi Square Test to see if there is an association
chiTestHairVsSex <- chisq.test(hairVsSex)
chiTestHairVsSex

chiTestHairVsEye <- chisq.test(hairVsEye)
chiTestHairVsEye

chiTestEyeVsSex <- chisq.test(eyeVsSex)
chiTestEyeVsSex

traitsCMHTest <- mantelhaen.test(HairEyeColor)
traitsCMHTest # Making sure something has association
```

From our tests we know the null hypothesis is that there is no association between variables and from the alternative that there is association. The $\alpha$ level for these tests are 0.05.

- We can see from the Hair and Eye variables that the p-value from this test is `r chiTestHairVsEye$p.value`. From this result, we can reject the null hypothesis and there is an association between hair and eye colors.

- We can see from the Hair and Sex variables that the p-value from this test is `r chiTestHairVsSex$p.value`. From this result, we can reject the null hypothesis and there is an association between hair color and a person's sex.

- We can see from the Sex and Eye variables that the p-value from this test is `r chiTestEyeVsSex$p.value`. From this result, we can fail to reject the null hypothesis and conclude that there is most likely *NOT* an association between a person's sex and eye color.

- Lastly, we can see from the Cochran-Mantel-Haenszel test that there was some kind of association with our variables when they were all together with the p-value of `r traitsCMHTest$p.value`. We can see from before that the hair color is associated with the other two variables which causes this result.

### Problem 2 {.tabset}

#### Part A

For each department, carry out a chi-square test of independence and report whether there is an association between gender and admission.
```{r}
# Load in the data
ucbData <- UCBAdmissions

# Do Dept. A test
deptA <- chisq.test(ucbData[ , , 1])
deptA

# Do Dept. B test
deptB <- chisq.test(ucbData[ , , 2])
deptB

# Do Dept. C test
deptC <- chisq.test(ucbData[ , , 3])
deptC

# Do Dept. D test
deptD <- chisq.test(ucbData[ , , 4])
deptD

# Do Dept. E test
deptE <- chisq.test(ucbData[ , , 5])
deptE

# Do Dept. F test
deptF <- chisq.test(ucbData[ , , 6])
deptF
```

From our tests we know the null hypothesis is that there is no association between variables and from the alternative that there is association. The $\alpha$ level for these tests are 0.05.

- From Dept. A, we get a p-value of `r deptA$p.value` which shows there is an association between gender and admissions.

- From Dept. B, we get a p-value of `r deptB$p.value` which shows there is *NOT* an association between gender and admissions.

- From Dept. C, we get a p-value of `r deptC$p.value` which shows there is *NOT* an association between gender and admissions.

- From Dept. D, we get a p-value of `r deptD$p.value` which shows there is *NOT* an association between gender and admissions.

- From Dept. E, we get a p-value of `r deptE$p.value` which shows there is *NOT* an association between gender and admissions.

- From Dept. F, we get a p-value of `r deptF$p.value` which shows there is *NOT* an association between gender and admissions.

**The only department that had this bias was department A.**

#### Part B

Ignoring department, carry out a single chi-square test of independence for the whole data and report whether there is an association between gender and admission.
```{r}
allDepts <- ucbData[ , , 1] + ucbData[ , , 2] + ucbData[ , , 3] + ucbData[ , , 4] + ucbData[ , , 5] + ucbData[ , , 6]
allDeptsTest <- chisq.test(allDepts)
allDeptsTest
```

From our test we know the null hypothesis is that there is no association between variables and from the alternative that there is association. The $\alpha$ level for these tests are 0.05. From combining all departments together, we get a p-value of `r allDeptsTest$p.value` which shows there is an association between gender and admissions.

#### Part C

Carry out a CMH chi-square test and report whether there is an association between gender and admission.
```{r}
cmhTest <- mantelhaen.test(ucbData)
cmhTest
```

From our test we know the null hypothesis is that there is no association between variables and from the alternative that there is association. The $\alpha$ level for these tests are 0.05. From combining all departments together, we get a p-value of `r cmhTest$p.value` which shows there is *NOT* an association between gender and admissions.

#### Part D

We have conflicting results in parts (a), (b), and (c). This is due to **Simpson's Paradox** since Department A shows there is association and has much more weight or amount of observations than the other departments grouped with it. Because the CMH Test and when looking at the departments individually, I believe that there is any association between gender and admissions, except for maybe Department A, as shown in part (a). 

#### Part E

Construct a summary matrix with success rates for male and female applicants in each department. Also calculate the overall success rate (i.e., ignoring department) of male and female candidates. From these numbers (without referring to statistical tests) what is your empirical conclusion --- do you  think there is gender bias in admissions? Why or why not?
```{r}
dept <- c("A", "B", "C", "D", "E", "F")
admit <- matrix(0, nrow = 6, ncol = 2)

for (i in 1:6){
  foo <- ucbData[,,i]
  admit[i, 1] = foo[1, 1] / (foo[1, 1] + foo[2, 1])
  admit[i, 2] = foo[1, 2] / (foo[1, 2] + foo[2, 2])
}

row.names(admit) <- dept
colnames(admit) <- c("Male", "Female")
round(admit, 3)
```

Personally just looking at the table, I believe that there is not any kind of gender bias present. The admission rates for each gender is really close, except for Department A. The reason for this department having differences is because there was about an 8 to 1 ratio of males applying to females. We might see this number regress towards the actual admissions rate if there were more female applicants. With the others so close, I believe there is no kind of gender bias present.

### Problem 3 {.tabset}

#### Part A

Load the `Groceries` transactions database from the `arules` package. Then we will answer some questions about it.
```{r}
library(arules)
data("Groceries")

Groceries # Here we can see that the number of transactions is 9,835 and the number of items is 169.

itemFrequencyPlot(Groceries, topN = 10, type="absolute") # Here we can see the top items that are being bought and here "Whole Milk" is the most bought one

percentReach20 <- sum(size(Groceries) >= 20) / length(size(Groceries)); percentReach20 # Gives us decimal form of groceries reaching at least 20 items in a transaction

meanTransaction <- mean(size(Groceries)); meanTransaction # Gives us mean number of items in a transaction
```

- First we can see that the number of transactions in this data set is `r nrow(Groceries)` and the number of items is `r ncol(Groceries)`.

- We can also see from our plot what the top items bought were. In this case *Whole Milk* was the most bought item.

- We can see that the percentage of transactions that involved at least 20 items was `r 100 * percentReach20`%. 

- Lastly, we can see that the average number of items in each transaction is `r meanTransaction`. 

#### Part B

Find all rules with support > 1% and confidence > 50%.
```{r}
rules <- apriori(Groceries, parameter = list(supp = 0.01, conf = 0.50))
rules # Shows us number of rules exist

inspect(rules) # Looks at all rules

# Find the highest confidence rule
highConfidence <- sort(rules, by = "confidence", decreasing = T)
inspect(highConfidence); inspect(highConfidence[1]) # Shows highest one
highConfidence_lhs <- "citrus fruit and root vegetables"
highConfidence_rhs <- "other vegetables"

# Find the highest support rule
highSupport <- sort(rules, by = "support", decreasing = T)
inspect(highSupport); inspect(highSupport[1]) # Shows highest one
highSupport_lhs <- "other vegetables and yogurt"
highSupport_rhs <- "whole milk"
```

- We found that there are `r length(rules)` rules with support > 1% and confidence > 50%.

- We can see that the rule with the highest confidence has on the left hand side `r highConfidence_lhs` and on the right hand side `r highConfidence_rhs`. We can see that it has a confidence of `r highConfidence@quality$confidence[1]`, a support of `r highConfidence@quality$support[1]`, and lift of `r highConfidence@quality$lift[1]`. The support is telling us that `r highConfidence@quality$support[1] * 100`% of customers purchase `r highConfidence_lhs`. The confidence is telling us that when `r highConfidence_lhs` is purchased, `r highConfidence_rhs` is shown up with it in the same transaction `r highConfidence@quality$confidence[1] * 100`% of the time. Lastly, the lift tells us that our confidence that they will buy `r highConfidence_rhs` increases by `r highConfidence@quality$lift[1] * 100`% when we know `r highConfidence_lhs` has been purchased.

- We can see that the rule with the highest support has on the left hand side `r highSupport_lhs` and on the right hand side `r highSupport_rhs`. We can see that it has a confidence of `r highSupport@quality$confidence[1]`, a support of `r highSupport@quality$support[1]`, and lift of `r highSupport@quality$lift[1]`. The support is telling us that `r highSupport@quality$support[1] * 100`% of customers purchase `r highSupport_lhs`. The confidence is telling us that when `r highSupport_lhs` is purchased, `r highSupport_rhs` is shown up with it in the same transaction `r highSupport@quality$confidence[1] * 100`% of the time. Lastly, the lift tells us that our confidence that they will buy `r highSupport_rhs` increases by `r highSupport@quality$lift[1] * 100`% when we know `r highSupport_lhs` has been purchased.

#### Part C

Which items do “whole milk” lead to? find all rules with support > 1%, confidence > 20%, and “whole milk” on the left-hand side. Report these rules.
```{r}
rulesMilkLeft <- apriori(Groceries, 
                         appearance = list(default = "rhs", 
                                           lhs = "whole milk"),
                 parameter = list(supp = 0.01, 
                                  conf = 0.20))
inspect(rulesMilkLeft)
```

Here from this table created, we can see that knowing that the customer has "whole milk" leads to `r length(rulesMilkLeft)` types of additional purchases. The rules are listed below.

1. Whole Milk leads to Yogurt
1. Whole Milk leads to Rolls/Buns
1. Whole Milk leads to Other Vegetables

#### Part D

Which items lead to “whole milk”? find all rules with support > 1%, confidence > 20%, and “whole milk” on the left-hand side. Report these rules.
```{r}
rulesMilkRight <- apriori(Groceries, 
                         appearance = list(default = "lhs", 
                                           rhs = "whole milk"),
                 parameter = list(supp = 0.01, 
                                  conf = 0.20))
inspect(rulesMilkRight)
```

Here from this table created, we can see that knowing that the customer has whole milk leads to `r length(rulesMilkRight)` types of additional purchases. Since there are so many, please look at the table above to see all the items that can lead to "whole milk".